Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
minor

\subsection{Permutations: rankings vs orderings}

@@

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model Algorithm}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} $r_1, r_2$: parameters controlling the learning rate 
    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign \{\}$
    \For{$i \assign 0$ \textbf{to} $\minit$}
    \State $\sigma \assign \text{generate uniformly at random}$
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor
    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}\Comment{Learning Step}
    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
   % \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     \State $\sigma \assign$ sample a permutation from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    % \Comment{Generate $\beta$ permutations}
    %\State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
    %\Statex \Comment{Select the farthest from $S$}
   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor 
    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

@@

\subsection{Rankings and orderings}
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

@@

%\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}
\bibliographystyle{ACM-Reference-Format-abbrv}
\bibliography{optbib/abbrev,optbib/authors,optbib/abbrevshort,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}

>>>>

\subsection{Permutations: rankings vs orderings}\label{sec:permutations}

@@

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model Algorithm}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} $r_1, r_2$: parameters controlling the learning rate 
    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign \{\}$
    \For{$i \assign 0$ \textbf{to} $\minit$}
    \State $\sigma \assign \text{generate uniformly at random}$
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor
    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S,\rho)$\label{line:uborda}\Comment{Learning Step}
    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
   % \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     \State $\sigma \assign$ sample a permutation from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    % \Comment{Generate $\beta$ permutations}
    %\State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
    %\Statex \Comment{Select the farthest from $S$}
   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor 
    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

@@

\subsection{Rankings and orderings}
In Section~\ref{sec:permutations} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of misinterpreting permutations. We remind the reader that statistical models (Borda or the \ken distance) are defined for rankings while standard literature in optimization considers orderings. This is the reason while in the optimization field we will usually find the Swap distance instead of the \ken distance. However, both are usually claimed to be the same, which is false: they are the same in particular cases, for example when  one of the permutations is the identity or an inversion.

@@

%\renewcommand{\doi}[1]{doi:\hspace{.16667em plus .08333em}\discretionary{}{}{}\href{https://doi.org/#1}{\urlstyle{rm}\nolinkurl{#1}}}

\bibliographystyle{ACM-Reference-Format-abbrv}
\bibliography{optbib/abbrev,optbib/authors,optbib/abbrevshort,optbib/journals,optbib/biblio,optbib/crossref}%,./mendeley}
